---
title: "COVID19_02"
author: "ym.deng"
date: "2023-07-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r }
library(tidyverse)
library(lubridate)
```

```{r}
case_us <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv")
death_us <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv")
case_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
death_global <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

```

**Pivot the dates and cases from rows to columns for calculation**
```{r}
case_global <- case_global %>%
  pivot_longer(cols = -c('Province/State', 'Country/Region', Lat, Long), 
               names_to = "date", 
               values_to = "cases") %>%
  select(-c(Lat, Long))

death_global <- death_global %>%
  pivot_longer(cols = -c('Province/State', 'Country/Region', Lat, Long),
               names_to = "date",
               values_to = "deaths") %>%
  select(-c(Lat, Long))
```

**Join the case and death into one table**
```{r}
global <- case_global %>%
  full_join(death_global) %>%
  rename(Country_Region = 'Country/Region',Province_State = 'Province/State') %>%
  mutate(date = mdy(date))

cat("**Filter out zero case**")
global <- global %>% filter(cases>0)


global <- global %>%
  unite("Combined_Key", c(Province_State, Country_Region), 
        sep=", ", 
        na.rm = TRUE, 
        remove = FALSE)

population_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"

population <- read_csv(population_url)

global <- global %>% 
  left_join(population, by = c("Province_State", "Country_Region")) %>%
  select(-c(UID, FIPS)) %>%
  select(Province_State, Country_Region, date, cases, deaths, Population)

```

**Top 10 countris with most cases**
```{r}
top_cases <- global %>%
  group_by(Country_Region) %>%
  summarise(Total_Cases = max(cases)) %>%
  top_n(10, Total_Cases) %>%
  arrange(desc(Total_Cases))

ggplot(top_cases, aes(x=reorder(Country_Region, Total_Cases), y=Total_Cases)) + 
  geom_bar(stat="identity", fill="steelblue") + 
  coord_flip() + 
  labs(title="Top 10 Countries with Most Cases", x="Country", y="Total Cases") +
  theme_minimal()

```

**Top 10 Countries with Most Deaths**
```{r}
top_deaths <- global %>%
  group_by(Country_Region) %>%
  summarise(Total_Deaths = max(deaths)) %>%
  top_n(10, Total_Deaths) %>%
  arrange(desc(Total_Deaths))

ggplot(top_deaths, aes(x=reorder(Country_Region, Total_Deaths), y=Total_Deaths)) + 
  geom_bar(stat="identity", fill="steelblue") + 
  coord_flip() + 
  labs(title="Top 10 Countries with Most Deaths", x="Country", y="Total Deaths") +
  theme_minimal()

```

**Global Cases Over Time**
```{r}
global_cases_time <- global %>%
  group_by(date) %>%
  summarise(Total_Cases = sum(cases))

ggplot(global_cases_time, aes(x=date, y=Total_Cases)) +
  geom_line() +
  labs(title="Global COVID-19 Cases Over Time", x="Date", y="Total Cases") +
  theme_minimal()
```

**Global Death Over Time**
```{r}
global_deaths_time <- global %>%
  group_by(date) %>%
  summarise(Total_Deaths = sum(deaths))

ggplot(global_deaths_time, aes(x=date, y=Total_Deaths)) +
  geom_line() +
  labs(title="Global COVID-19 Deaths Over Time", x="Date", y="Total Deaths") +
  theme_minimal()
```

**Global Case Fatality Rate**
```{r}
global <- global %>%
  group_by(date) %>%
  summarise(Total_Cases = sum(cases),
            Total_Deaths = sum(deaths)) %>%
  mutate(Case_Fatality_Rate = Total_Deaths / Total_Cases)

ggplot(global, aes(x=date, y=Case_Fatality_Rate)) +
  geom_line() +
  labs(title="Global COVID-19 Case Fatality Rate Over Time", x="Date", y="Case Fatality Rate") +
  theme_minimal()
```


**Us Data Cleaning**
```{r}
case_us <- case_us %>%
  pivot_longer(cols = -c(UID:Combined_Key), 
               names_to = "date",
               values_to = "cases") %>%
  select(Admin2:cases) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

death_us <- death_us %>%
  pivot_longer(cols = -c(UID:Population),
               names_to = "date",
               values_to = "deaths") %>%
  select(Admin2:deaths) %>%
  mutate(date = mdy(date)) %>%
  select(-c(Lat, Long_))

```

```{r}
us <- case_us %>%
  full_join(death_us)
```

***death rate of us*** 
```{r}
us_by_state <- us %>%
  group_by(Province_State, Country_Region, date) %>%
  summarise(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  select(Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()

us_totals <- us_by_state %>%
  group_by(Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  select(Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()

```


***us total and city death and cases time series***
```{r}
us_totals %>%
  filter(cases > 0 ) %>%
  ggplot(aes(x = date, y = cases)) + 
  geom_line(aes(color = "cases")) + 
  geom_point(aes(color = "cases")) + 
  geom_line(aes(y = deaths, color = "deaths")) + 
  geom_point(aes(y = deaths, color = "deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "botton", axis.text.x = element_text(angle = 90)) + 
  labs(title = "COVID19 in US", y = NULL )

us_by_state %>%
  filter(Province_State == "New York") %>%
  filter(cases > 0) %>%
  ggplot(aes(x = date, y = cases)) + 
  geom_line(aes(color = "cases")) + 
  geom_point(aes(color = "cases")) + 
  geom_line(aes(y = deaths, color = "deaths")) + 
  geom_point(aes(y = deaths, color = "deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "botton", axis.text.x = element_text(angle = 90)) + 
  labs(title = "COVID19 in US", y = NULL )

max(us_totals$date)


```


***US new_case and new death time series***
```{r}
us_by_state <- us_by_state %>%
  mutate(new_cases = cases - lag(cases), 
         new_deaths = deaths - lag(deaths))

us_totals <- us_totals %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths))

us_totals %>%
  filter(cases > 0 ) %>%
  ggplot(aes(x = date, y = new_cases)) + 
  geom_line(aes(color = "new_cases")) + 
  geom_point(aes(color = "new_cases")) + 
  geom_line(aes(y = new_deaths, color = "new_deaths")) + 
  geom_point(aes(y = new_deaths, color = "new_deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "botton", axis.text.x = element_text(angle = 90)) + 
  labs(title = "COVID19 in US", y = NULL )


us_state_totals <- us_by_state %>%
  group_by(Province_State) %>%
  summarize(deaths = max(deaths), cases = max(cases),
            population = max(Population),
            cases_per_thou = 1000 * cases / population, 
            deaths_per_thou = 1000 * deaths / population) %>%
  filter(cases > 0, population > 0 )
```


***us state top death rate*** 
```{r}
us_state_totals %>%
  slice_min(deaths_per_thou, n=10) %>% 
  select(deaths_per_thou, cases_per_thou, everything())

us_state_totals %>%
  slice_max(deaths_per_thou, n=10) %>% 
  select(deaths_per_thou, cases_per_thou, everything())

```


*** Model Prediction of Deaths to Cases*** 
```{r}
mod <- lm(deaths_per_thou ~ cases_per_thou, data = us_state_totals)
summary(mod)


us_tot_w_pred <- us_state_totals %>% mutate(pred = predict(mod))

us_tot_w_pred %>% ggplot() + 
  geom_point(aes(x = cases_per_thou, y = deaths_per_thou), color = "blue") +
  geom_point(aes(x = cases_per_thou, y = pred), color = "red")

```
